\bf Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-20 appendix.qmd
Updated Figure 4

```{r}
#| out-width: 50%
#| eval: false

result_umap |>
  filter(min.dist == 0.5) |>
  ggplot(aes(
           x = as.numeric(n_neighbors),
           y = number_dubious_cells
         )) +
    geom_line() +
    geom_point() +
    labs(x = "n_neighbors",
         y = "# of dubious cells") +
    theme_minimal()
```

```{r}
#| label: read-pbmc-umap-layouts
#| eval: false
# Read a variety of different NLDR representations of PBMC
# and plot them on same aspect ratio
umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
umap_pbmc1 <- as_tibble(umap_pbmc1)
names(umap_pbmc1) <- c("UMAP1", "UMAP2")

@@

```{r}
#| label: combine-data-pbmc-umap
#| eval: false

@@

```{r}
#| label: error-comp-pbmc-umap
#| eval: false

@@

```{r}
#| label: fig-pbmc-mse-umap
#| eval: false
#| fig-pos: H
#| out-height: 53%
#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the top plot and in the scatterplots represents UMAP layout (a, b). Of the two, layout b is optimal across all binwidths making it the best choice."

>>>>

```{r}
#| label: fig-n_neighbors-dubious_cells
#| out-width: 50%
#| fig-pos: H
#| eval: false

result_umap |>
  filter(min.dist == 0.5) |>
  filter(n_neighbors < 240) |>
  ggplot(aes(
           x = as.numeric(n_neighbors),
           y = number_dubious_cells
         )) +
    geom_line() +
    geom_point() +
    labs(x = "n_neighbors",
         y = "Number of dubious cells") +
    theme_minimal()
```

```{r}
#| label: read-pbmc-umap-layouts
# Read a variety of different NLDR representations of PBMC
# and plot them on same aspect ratio
umap_pbmc1 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.3.rds")
umap_pbmc1 <- as_tibble(umap_pbmc1)
names(umap_pbmc1) <- c("UMAP1", "UMAP2")

@@

```{r}
#| label: combine-data-pbmc-umap

@@

```{r}
#| label: error-comp-pbmc-umap

@@

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-width: 100%
#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the top plot and in the scatterplots represents UMAP layout (a, b). Of the two, layout b is optimal across all binwidths making it the best choice."

### 2025-04-20 appendix.tex
Updated Figure 3
Added the corrected .bib
Updated Figure 4

\includegraphics[width=1\linewidth,height=0.3\textheight]{appendix_files/figure-pdf/fig-tri-param-1.pdf}

@@

\textbf{?@fig-pbmc-mse-umap} shows layout from the original published
article along with that suggested by scDEED to be more accurate. The log
MSE vs binwidth plot illustrates that our approach would suggest that
scDEED is correct here, that layout b is more accurately reflecting the
cluster structure in the data.

  \bibliography{bibliography.bib}

>>>>

\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-tri-param-1.pdf}

@@

Figure~\ref{fig-pbmc-mse-umap} shows layout from the original published
article along with that suggested by scDEED to be more accurate. The log
MSE vs binwidth plot illustrates that our approach would suggest that
scDEED is correct here, that layout b is more accurately reflecting the
cluster structure in the data.

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}

\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
with different hyper-parameter settings (n\_neighbors: \(30\),
min\_dist: \(0.3\) (red); n\_neighbors: \(80\), min\_dist: \(0.5\)
(blue)) on the PBMC data is the better representation using MSE for
varying binwidth (\(a_1\)). Colour used for the lines and points in the
top plot and in the scatterplots represents UMAP layout (a, b). Of the
two, layout b is optimal across all binwidths making it the best
choice.}

  \bibliography{paper.bib}
